---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Loading librarys needed
```{r,warning=FALSE,message=FALSE}
library(dplyr)
```
### load and prepare the data
```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
head(activity)
```


## What is mean total number of steps taken per day?
### First calculate the sums of steps per day
```{r}
sum_steps <- activity %>% filter(!is.na(steps)) %>% group_by(date) %>% 
    summarise(tot_steps = sum(steps))
arrange(sum_steps, desc(tot_steps))
```

### create histogram
```{r}
hist(sum_steps$tot_steps, breaks = 25, main = "Histogram of total number of steps taken per day", xlab = "steps")
```

### calculate the mean
```{r}
mean_steps <- mean(sum_steps$tot_steps)
median_steps <- median(sum_steps$tot_steps)
```
mean = `r mean_steps`   
median = `r median_steps`

## What is the average daily activity pattern?
### First calculate the sums of steps per interval
```{r}
steps_interval <- activity %>% filter(!is.na(steps)) %>% group_by(interval) %>% 
    summarise(mean_steps_int = mean(steps))
arrange(steps_interval, desc(mean_steps_int))
```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
