---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Loading librarys needed
```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
```
### load and prepare the data
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
# download file if not already in Wd
if(!file.exists("repdata%2Fdata%2Factivity.zip"))
{
  download.file(fileUrl, "activity.zip")
}
# Unzip if necessary
if (!file.exists("activity.csv")) 
{ 
  unzip("activity.zip") 
}

activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
head(activity)
```


## What is mean total number of steps taken per day?
### First calculate the sums of steps per day
```{r}
sum_steps <- activity %>% filter(!is.na(steps)) %>% group_by(date) %>% 
    summarise(tot_steps = sum(steps))
arrange(sum_steps, desc(tot_steps))
```

### create histogram
```{r}
hist(sum_steps$tot_steps, breaks = 25, main = "Histogram of total number of steps taken per day", xlab = "steps")
```

### calculate the mean
```{r}
mean_steps <- mean(sum_steps$tot_steps)
median_steps <- median(sum_steps$tot_steps)
```
mean = `r mean_steps`   
median = `r median_steps`

## What is the average daily activity pattern?
### First calculate the sums of steps per interval
```{r}
steps_interval <- activity %>% filter(!is.na(steps)) %>% group_by(interval) %>% 
    summarise(mean_steps_int = mean(steps))
arrange(steps_interval, desc(mean_steps_int))
```

### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
with(steps_interval, plot(interval, mean_steps_int, type = "l", xlab = "interval", 
                          ylab= "Average number of Steps", main = "Average number of steps during the day"))
```

### Determine the 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps.
```{r}
max_interval <- steps_interval %>% filter(mean_steps_int == max(mean_steps_int)) %>% select(interval)
```
interval that contains the maximum number of steps on average = `r max_interval$interval` 

## Imputing missing values
### Calculate and report the total number of missing values
```{r}
sum_na <-sum(is.na(activity$steps))
```
total number of missing values = `r sum_na` 

### Impute missing values with the mean for each interval
```{r}
activity_nw <- activity
activity_nw$steps[is.na(activity_nw$steps)] <- tapply(activity_nw$steps, activity_nw$interval, mean, na.rm=TRUE, simplify=T)
```

### make histogram and calculate the mean and median per day
```{r}
sum_steps_nw <- activity_nw %>% filter(!is.na(steps)) %>% group_by(date) %>% 
    summarise(tot_steps = sum(steps))
hist(sum_steps_nw$tot_steps, breaks = 25, main = "Histogram of total number of steps taken per day", xlab = "steps")
```

### calculate the mean
```{r}
mean_steps_nw <- mean(sum_steps_nw$tot_steps)
median_steps_nw <- median(sum_steps_nw$tot_steps)
```
mean = `r mean_steps_nw`   
median = `r median_steps_nw`  
imputing data has no effect on the mean and a small change in median (is equal to mean now)

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity_nw <- mutate(activity_nw, 
                      wd_ind = as.factor(case_when(weekdays(activity_nw$date) == "Sunday" |
                                                   weekdays(activity_nw$date) == "Saturday" ~ "weekend"
                                                   , TRUE ~ "weekday")))
g <- ggplot(activity_nw, aes(interval, steps))
